---
title: "final_project.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(stringr)

```

```{r}
# Denormalize, combining results, phases (a.k.a. heats), events
detailed_results <- fina_join()
```

## Filter to separate competitions (same named competitions)
## 99 competitions 
```{r}
comps <- detailed_results %>% group_by(competition) %>% summarise(n = n())
sepdetailed_results <- detailed_results %>% separate(col = competition, 
                                                     into = c("competitionname", "year"),
                                                     sep = -4) 
sepdetailed_results

word(competitionname, start = 205, end = 259, sep = fixed)

test1 <- finaldetailed_results %>% group_by(compname) %>% summarise(n = n())
sepdetailed_results %>% group_by(series) %>% summarise (n = n())
```

```{r}

```

## Filter by competition
```{r}
test1 <- detailed_results %>% filter(series == "Olympic Games")

```

## Filter by event (buttons?)
```{r}
test2 <- test1 %>% filter(style == "Medley")
```

## Filter by length of the event (buttons?)
```{r}
test3 <- test2 %>% filter(distance == "200")
```

## Filter by whether the even is a relay
```{r}
test4 <- test3 %>% filter(relay == FALSE)
```

## Filter by the year
```{r}
test5 <- test4 %>% filter(year == 2008)
```

```{r}
test5
```

## Filter by gender
```{r}
test6 <- test5 %>% filter(gender == "Men")

test6 %>% group_by(phase_label) %>% summarise(n = n())
test7 <- test6 %>% filter(phase_label == "Final")

```

## Make a table of the three winners, their country, and their time
```{r}
test7 <- test6 %>% filter(phase_label == "Final")
test8 <- test7 %>% filter(rank == 1 | rank == 2 | rank == 3) %>% select(rank, ioc_code, family_name, first_name, time)
test8
```

## Olympic Games worked

```{r}
test <- detailed_results %>% filter(series == "Olympic Games") %>%
  filter(gender == "Men") %>%
  filter(relay == FALSE) %>%
  filter(distance == 200) %>%
  filter(style == "Breaststroke") %>%
  group_by(year) %>%
  summarise(averagetime = mean(time, na.rm = TRUE),
                se = sd(time, na.rm = TRUE)/sqrt(100),
                l_se = averagetime - se,
                u_se = averagetime + se)
```


## graph the change in average time
```{r}
library(plotly)
x <- ggplot(data = test, aes(x = year, y = averagetime)) +
  geom_point() +
  geom_errorbar(aes(ymin = l_se, ymax = u_se)) +
  labs(x = "Year", y = "Average Time", title = "How The Average Swim Time Has Changed Over The Years")

ggplotly(x)
```

## years in each series
```{r}
w <- detailed_results %>% filter(series == "Youth Olympic Games") %>% group_by(year) %>% summarise(n = n())
detailed_results %>% group_by(series) %>% summarise(n = n())
```

## map wins across each country for each year
```{r}
library(maps)
library(tidyverse)
detailed_results
library(countrycode)
countrycode(detailed_results$ioc_code, "ioc", "country.name")
## ask how to fix the NA variables
world_df <- ggplot2::map_data("world")
ggplot(data = world_df, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_map(projection = "gilber", xlim = c(-180, 180)) +
  theme_void()
## scale_fill_distiller(direction = 1)
results_full <- left_join(codelist_panel, detailed_results, by = c("ioc" = "ioc_code"))
codelist_panel
test <- detailed_results %>% filter(series == "Olympic Games")
results_half <- left_join(codelist_panel, detailed_results, by = c("ioc" = "ioc_code"))
```

### Side Notes
## Possibly be able to change the options for inputs depending on the year selected and other? inputs
## Look into the World Cup rankings because there are two groups of three top winners for each (data issue?)
## Make a map that shows the wins across each country for each year
## Want to add a visualization that explores times each year (possibly be able to add a time and see what rank that would get you)

## won't let me combine data sets because too big ?

## Look at reaction time vs results

```{r}
test <- detailed_results %>% filter(series == "Olympic Games") %>%
  filter(gender == "Men") %>%
  filter(relay == FALSE) %>%
  filter(distance == 100) %>%
  filter(style == "Butterfly") %>%
  filter(year == 2008) %>%
  filter(phase_label == "Final")

ggplot(data = test, aes(x = time, y = rt)) +
  geom_point()

test2 <- detailed_results %>% filter(series == "Olympic Games") %>%
  filter(gender == "Men") %>%
  filter(relay == TRUE) %>%
  filter(distance == 400) %>%
  filter(style == "Freestyle") %>%
  filter(year == 2008)
View(test2)

mod1 <- lm(time~gender + relay + distance + style + year + series + rt, data = detailed_results)

```

## Make something that when the time is put in it will show what place you would've gotten that year

```{r}
rbind(filtered_results, c())
```

```{r}
test2 <- detailed_results %>% filter(series == "Olympic Games") %>%
  filter(gender == "Men") %>%
  filter(relay == FALSE) %>%
  filter(distance == 400) %>%
  filter(style == "Freestyle") %>%
  filter(year == 2008) %>%
  filter(phase_label == "Final")
test2[is.na(test2)] <- 0
ggplot(data = test2, aes(x = timebehind, y = first_name)) +
  geom_col()
```

## Make something that when the time is put in it will show what place you would've gotten that year


### Lists top 8 finishers and the time behind from the first finisher they were
## add something that represents the country to this graph.
```{r}
test2 <- detailed_results %>% filter(series == "Olympic Games") %>%
  filter(gender == "Men") %>%
  filter(relay == FALSE) %>%
  filter(distance == 400) %>%
  filter(style == "Freestyle") %>%
  filter(year == 2008) %>%
  filter(phase_label == "Final") %>%
  arrange(desc(phase_label)) %>%
  slice(1:8) %>%
  mutate(timebehind = case_when(is.na(time_behind) ~ "0", 
          !is.na(time_behind) ~ time_behind)) %>%
  mutate(name_ordered = fct_reorder(.f = family_name, .x = timebehind))
View(test2)
# get na value out

library(plotly)

ggplot(data = test2, aes(x = name_ordered, y = timebehind)) +
  geom_col() +
  coord_flip() +
  labs(x = "Time Behind",
       y = "Swimmer")
```

## medals by country

```{r}
test4 <- detailed_results %>% filter(series == "Olympic Games") %>%
  filter(phase_label == "Final") %>%
  filter(rank == 1 | rank == 2 | rank == 3) %>%
  filter(year == 2008) %>%
  group_by(ioc_code, rank) %>%
  summarise(medals = n()) %>%
  pivot_wider(names_from = rank,
              values_from = medals) %>%
  mutate(gold = as.numeric(`1`),
                 silver = as.numeric(`2`),
                 bronze = as.numeric(`3`)) %>% 
  mutate(Gold = case_when(is.na(gold) ~ 0, !is.na(gold) ~ gold),
         Silver = case_when(is.na(silver) ~ 0, !is.na(silver) ~ silver),
         Bronze = case_when(is.na(bronze) ~ 0, !is.na(bronze) ~ bronze)) %>%
  mutate(totalmedals = Gold + Silver + Bronze) %>%
  select(ioc_code, Gold, Silver, Bronze, totalmedals)
  

test4 %>% mutate(gold = as.numeric(gold),
                 silver = as.numeric(`2`),
                 bronze = as.numeric(`3`)) %>% 
  mutate(Gold = case_when(is.na(gold) ~ 0, !is.na(gold) ~ gold),
         Silver = case_when(is.na(silver) ~ 0, !is.na(silver) ~ silver),
         Bronze = case_when(is.na(bronze) ~ 0, !is.na(bronze) ~ bronze))


test4$gold
test4[is.na(test4)] = 0
test4
library(knitr)
kable(test4, col.names = c("Country", "Gold Medals", "Silver Medals", "Bronze Medals", "Total Medals"))
library(pander)
pander(test4)
```

```{R}
library(plotly)
x <- ggplot(data = test, aes(x = year, y = averagetime)) +
  geom_point() +
  geom_errorbar(aes(ymin = l_se, ymax = u_se)) +
  labs(x = "Year", y = "Average Time", title = "How The Average Swim Time Has Changed Over The Years")

ggplotly(x)
```

```{r}
detailed_results %>% filter(series == "Youth Olympics") %>%
      filter(phase_label == "Final") %>%
      filter(rank == 1 | rank == 2 | rank == 3) %>%
      filter(year == 2006) %>%
      group_by(ioc_code, rank) %>%
      summarise(medals = n()) %>%
      pivot_wider(names_from = rank,
                  values_from = medals)

detailed_results %>% filter(series == "Youth Olympics") 
detailed_results %>% group_by(series) %>%
  summarise(n = n())
```
